A Study of Residual Adapters for Multi-Domain Neural Machine Translation

### 2020-08-14 2020/wmt_20/emnlp2020.tex

@@

In this context, our main contribution is a thorough experimental study of the use of residual adapters in a multi-domain translation adaptation. We notably explore ways to adjust and/or regularize adapter modules to handle situations where the adaptation data is very small. We also propose and contrast two new variants of the residual architecture: in the first one (highway residual adapters), adaptation still affects each layer of the architecture, but its effect is delayed till the last layer, thus making the architecture more modular and adaptive; in the second one (gated residual adapters), we explore ways to improve the performance in the face of train-test data mismatch. We experiment with two langage pair and report results that demonstrate the flexibility and effectiveness of these architectures. 
\fyDone{One bit of a conclusion here}\fyFuture{Build a proper training scenario for these two conditions}

@@

\subsection{Architecture \label{ssec:architecture}}
\fyTodo{More contexts and notations from the transformer}\fyTodo{Encoder / decoder layers}
For $i^{\text{th}}$layer of the encoder or the decoder, let $h^i$ be the output of the layer, of dimension $d$. The $i^{\text{th}}$ residual adapter modifies $h^i$ by applying several transformations as follows:\fyDone{Use align env}
\begin{align*}
  h^{i}_1 &= \mathbf{W}_{db}^{i}h^{i} + b^i_{1} \\
  h^{i}_2 &= \mathbf{RELU}(h_1^{i}) \\
  h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + b^i_{2} \\
  \bar{h}^{i} &= h^{i}_3 + h^i.
\end{align*}
Matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, b is a bottleneck dimension. For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the set of parameters for the $i^{th}$ adapters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}

The "adapted" context vector $\bar{h}^i_{ 1\leq i \leq n-1}$, where n is the number of layers, is the input of the $(i+1)^{\text{th}}$ \fyDone{Self attention ?} layer, $\bar{h}^n$ is passed to the decoder if it belongs to the encoder side, or the input of output layer if it belongs to the decoder side. In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder.\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter} 
\fyTodo{Many question opens: how to train ? how many adapters ? which size ? etc}

\subsection{Ablation study on position and number of residual adapters \label{ssec:ablatation}}
\fyTodo{Move this}
In this section, we would like to analyze the impact of position and number of residual adapters involved in the adapted model. In fact, the importance of hidden layers varies with respect to their position in the model. It is conjectured that the higher layers represent more global patterns such as semantic while the lower layers represent more local patterns such as syntactic\fyTodo{Missing reference here}. The domain shift in local patterns and global patterns has not yet studied. In this paper, we do not intend to study this aspect of domain adaptation problem. We would like to give an extensive comparison between domain adaptation in different levels in NMT model.\fyTodo{Fix style here} Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 which covers a wide range in the hierarchy of attention layers.

Second, we would like to observe how much the number of residual adapters affects the performance of the adapted model. By the limit of computation resource, we limit our study to the case of 3 adapters at positions: 2,4,6. By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting) in the encoder or the decoder.\fyTodo{We have 12 layers, have we not ?}

\subsection{Multi-task training\label{ssec:multitask}}
In the experiments of \cite{Caruana97multitask}, the authors proposed an architecture suitable for multi-task learning. The corresponding network comprises a ground layer, which is shared between tasks and is followed by task-related layers in the next level. The network is then trained from scratch by fueling task-related examples in a batch and updating all parameters (shared and task-specific) at the same time. The multi-domain context is similar to multi-task context if we consider a domain as a task\cite{Dredze08}. Therefore, we can also use multi-task training to optimize the parameters of NMT model and the parameters of the residual adapters at once from beginning. To assess the efficacy of multi-task training in multi-domain context, we would like to compare the finetuning approach with multi-task training approach. \fyTodo{Change the training regime}

\subsection{Regularization of residual adapters \label{ssec:reg}}
Fine-tuning on small corpora can still lead to overfitted model even when the fine-tuning is applied to a few residual adapters. To avoid this situation, we have several options including: reduce the size of the adapter module, apply weight decay to the adapter during adaptation, or apply layer regularization to the adapter modules. The first option is not flexible \fyTodo{compatible ???} if we consider the probable growth of in-domain data. The second and third choices are preferable to regularize a large adapter which will scale if the in-domain corpora extends.\fyTodo{Style}

Weight decay applies a penalization on the weights of the adapters.\fyDone{The second summation also runs over $x,y$ ? I think not}
\begin{equation}
  \begin{split}
    \bar{L} & = \mathop{\sum}_{x,y} - \log P(y|x) \\
    & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{\operatorname{ADAP}^{(i)}}})
  \end{split}
\end{equation}

Layer regularization applies a penalization on the output of the adapters.
\begin{equation}
  \begin{split}
    \bar{L} & = \mathop{\sum}_{x,y} (-\log(P(y|x)) \\
    & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\operatorname{ADAP}^{(i)}(h_i(x,y))})
  \end{split}
\end{equation}

\subsection{Highway Residual Adapters \label{ssec:highway}}
\fyTodo{Mettre mieux en valeur} \fyTodo{Meilleur entrainement du fait des connections highway ?}
Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers. We graphically demonstrate this version in Figure~\ref{fig:hrl-architecture}. The benefit of this variant is that we can manipulate all values of the adapters at once.

\subsection{Gated Residual Adaptaters \label{ssec:wada}}
\mpTodo{Formalizing problem, network design, training algorithm}
Residual adapters allow us to adapt NMT models to a new domain without altering the pretrained parameters. However, we still have to decide in which domain a sentence is translated to chose a suitable adapter.\fyTodo{Lack of context - domain errors} Therefore, we face to another problem which is the error of domain prediction. Choosing wrong adapter to translate a sentence can be dramatically bad because a specialized model usually has bad performance for out-of-domain instances \cite{McCloskey89catastrophic}. However, we still have another option that is to use the pretrained generic model without adaptation. Therefore, we could aim to achieve a performance at least as good as the generic model. We would like to present an adaptive version of residual adapter below.
\fyTodo{Consistency of notations wrt section 2.1}
We consider a generic encoder-decoder architecture for a multi-domain sequence to sequence learning. Let $h \in \R^d$ the output of the encoder, \cite{Bapna19simple} proposes to use an adapted representation for domain $k$ defined as $h' = h + a_k(h)$. This means that all words in all sentences in domain $k$ will use the adapter module represented by $a_k$. In a multilayer version with highway connections (see Figure~\ref{fig:hrl-architecture}), $a_k(h) = \sum_{l} a_{kl}(h^l)$ where $h^l$ is the distributed representation of the sequence at the $l^{th}$ layer.
\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{fig/highway_residual}
  \caption{Highway multi-domain network with residual layers}
\label{fig:hrl-architecture}
\end{figure}

A "gated" version uses $h' = h + a_k(h) \odot{} z_k(h)$\footnote{Or $a_k(h) = \sum_{l} a_{kl}(h)$, where the summation runs over layers, see Figure~\ref{fig:hrl-architecture}} where $z_k$ is a function of $h$ taking values in $[0,1]$. More precisely, $h$ and $h'$ are sequences of context vectors and the combination is performed element-wise, yielding:
\begin{equation}
   \forall t \in [1 \dots{} T], h'(w_t) = h(w_t) + a_k(h(w_t)) \odot{} z_k(h(w_t)). \label{eq:gated-residual}
\end{equation}

In Word Level Adaptive Domain Adaptation (WADA), $z_k(h(w))$ is designed to reflect the "topicality" of  word $w$ is in domain $k$: the more likely $w$ is in domain $k$, the larger $z_k(h(w))$ should be. Word-level adaptive domain adaptation aims to reduce the variation caused by $a_k$ to the adapted context vector $h'$ (compared to $h$) for words that are not typical of domain $k$. By reusing the learned generic representation for non-topical words, we can bound the risk of poor predictions in case of out-of-domain words by the risk of the learned generic model.

\subsubsection{The training process \label{sssec:train}}
The training process comprises three main steps:
\begin{itemize}
	\item Pretraining a generic model with a mixed corpora comprising data from multiple domains.
	\item Training a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen. This model computes the posterior domain probability $p(k|h(w_t))$ for each word $w_t$ based on the representation computed by the last layer.
	\item Training the parameters of adapters with in-domain data separately for each domain while freezing the parameters of the generic model and of the domain classifiers.
\end{itemize}
During inference, $z_k$ is used to regulate the strength of the adapter module as suggested in equation~\ref{eq:gated-residual}.

@@

\begin{table*}[htbp]
  \centering
  \begin{tabular}{ |lllllll|} %*{4}{|r|}}
    \hline
    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
    \domain{bank} & \domain{eco} & \domain{med} & \domain{gouv} & \domain{news} & \domain{tourism} & \domain{other} \\
    \hline
    4(0.00022) & 2857(0.15) & 347(0.018) & 1828(0.095) & 3696(0.19) & 7(0.00039) & 10472771(0.54) \\
    \hline
  \end{tabular}
\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture. \domain{news} is the largest domain, containing almost 19\% of the sentences, while \domain{bank} is the smallest, with only 0.02\% of the data.}
\label{tab:Corpora-en-de}
\end{table*}

@@

For all models, we set the embeddings size and the hidden layers size to~512. Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells. The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024. 
% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
Training use a batch size of~12288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}

\subsection{Varying the positions and number of residual adapters}

Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} for En-De. In most cases, the performance increases with respect to the number of residual adapters used in architecture. The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level. However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?}

\begin{table*}[htbp]
  \centering
  \begin{tabular}{|l|l|} \hline
    Model & params \\ \hline 
    \system{Transformer}  & 65 \\
   \system{Residual Adapter} & 1 \\
     \hline
  \end{tabular}
  \caption{Number of parameters ($\times 10^3$)}
  \label{tab:params}
\end{table*}

\begin{table*}
  \centering
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}} 
    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8} & \SB{53.8} \\
    \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2} & \SB{53.3} \\ 
    \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\ 
    \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 \\
    \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 \\
    \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 \\
   \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 \\
     \hline
  \end{tabular}
  \caption{Translation performance of various finetuned systems (En:Fr). We report BLEU scores for each domain, as well as averages.}\fyTodo{Boldface ?}
  \label{tab:performance-en-fr}
\end{table*}

@@

\fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ? It helps for small domain - domain specific regularization ??}

\subsection{Regularization of fine-tuning}
In the translation from English into German, we find that domains \domain{tourism}, \domain{ecb} are extremely small and account for a very small percentage of the training data, with respective proportions of 0.039\% and 0.022\%. Fine-tuning on these domains can lead to serious overfitting. We assess two well-known regularization techniques for adapter modules, that could help avoid the overfitting issues: weight decay and layer regularization. Results in Table~\ref{tab:performance-en-de-reg} show that weight decay (\system{FT-Block-WD}) outperforms the more basic \system{FT-Block}. \fyTodo{How is the weight decay parameter set ?}

\begin{table*}[htbp]
  \centering
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}} 
    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
   \system{FT-Block}     & 37.3	& 57.93 &	53.91 &	33.79 &	46.69 &	90.17 & 53.3 \\ 
   \system{FT-Block-WD}     & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 \\
   \system{FT-Block-LR}     & 37.45 & 56.09 & 51.84 & 33.29 & 45.02 & 89.7 & 52.2 \\
     \hline
  \end{tabular}
  \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
  \label{tab:performance-en-fr-reg}
\end{table*}

@@

\subsection{Multi-task learning}
To assess the effectiveness of multi-task learning in training both the NMT model and the residual adapters at once from scratch, we contrast our approach with several proposal from the literature on multi-domain MT including:

>>>>

\iffalse   % removes draft highlighting for final version
\renewcommand{\draftAdd}[2]{#1}
\fi

@@

In this context, our main contribution is a thorough experimental study of the use of residual adapters in a multi-domain translation adaptation. We notably explore ways to adjust and/or regularize adapter modules to handle situations where the adaptation data is very small. We also propose and contrast two new variants of the residual architecture: in the first one (highway residual adapters), adaptation still affects each layer of the architecture, but its effect is delayed till the last layer, thus making the architecture more modular and adaptive; in the second one (gated residual adapters), we explore ways to improve the performance in the face of train-test data mismatch. We experiment with two language pairs and report results that demonstrate the flexibility and effectiveness of these architectures. 
\fyDone{One bit of a conclusion here}\fyFuture{Build a proper training scenario for these two conditions}

@@

\subsection{Architecture \label{ssec:architecture}}
\fyTodo{More contexts and notations from the transformer}\fyTodo{Encoder / decoder layers}
For $i^{\text{th}}$layer of the encoder or the decoder, let $h^i$ be the output of the layer, of dimension $d$. The $i^{\text{th}}$ residual adapter modifies $h^i$ by applying several transformations as follows:\fyDone{Use align env}
\begin{align*}
  h^{i}_1 &= \mathbf{W}_{db}^{i}h^{i} + b^i_{1} \\
  h^{i}_2 &= \mathbf{RELU}(h_1^{i}) \\
  h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + b^i_{2} \\
  \bar{h}^{i} &= h^{i}_3 + h^i.
\end{align*}
Matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, b is a bottleneck dimension. For the sake of brevity, we denote $h^i_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the set of parameters for the $i^{\text{th}}$ adapters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}

The "adapted" hidden vectors $\bar{h}^i_{ 1\leq i \leq n-1}$, where n is the number of layers, is the input of the $(i+1)^{\text{th}}$ \fyDone{Self attention ?} layer, $\bar{h}^n$ is passed to the decoder if it belongs to the encoder side, or the input of output layer if it belongs to the decoder side. In our architecture, there are 12 residual adapters, one for each of the six attention layers of the encoder and similarly for the decoder.\footnote{In the decoder, the stack of self-attention layers and cross encoder-decoder attention only counts as one attention layer and serves only one residual adapter} 
\fyTodo{Many question opens: how to train ? how many adapters ? which size ? etc}

The size of available corpora for each domain is extremely varying. For example, the corpus of domain \domain{tourism} in En-De contains only around $7\times10^3$ sentence pairs while the corpus of domain \domain{news} in En-De contains approximately $3\times10^6$ sentence pairs. For extremely small domains, we could economize the number of parameters by reducing the number of residual adapters in the architecture. While only a limited number of adapters are incorporated to the NMT model, which positions in the model are more important in the domain adaptation task? We would like to analyze the impact of position and number of residual adapters involved in the adapted model. It is conjectured that the higher layers represent more global patterns such as semantic while the lower layers represent more local patterns such as syntactic \cite{raganato18analysis}\fyTodo{Missing reference here}. The domain shift in local patterns and global patterns has not yet studied. In this paper, we do not intend to study this aspect of the domain adaptation problem. We would like to give an extensive comparison between domain adaptation in different levels in NMT model.

Fine-tuning on small corpora can still lead to an overfitted model even when the fine-tuning is applied to a few residual adapters. To avoid this situation, we have several options including: reducing the size of the adapter module, applying weight decay or layer regularization to the adapter. The first option is not flexible \fyTodo{compatible ???} if we consider the probable growth of in-domain data. The second and third choices are preferable to regularize a large adapter which will scale if the in-domain corpora extend .\fyTodo{Style}

Weight decay applies a penalization on the weights of the adapters.\fyDone{The second summation also runs over $x,y$ ? I think not}
\begin{equation}
  \begin{split}
    \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y}( - \log(P(y|x))) \\
    & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{\operatorname{ADAP}^{(i)}}}
  \end{split}
\end{equation}
Where $\theta_{\operatorname{ADAP}^{(i)}}$ are parameters of $i^{\text{th}}$ adapter

Layer regularization applies a penalization on the output of the adapters.
\begin{equation}
  \begin{split}
    \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y} (-\log(P(y|x)) \\
    & + \lambda * \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\operatorname{ADAP}^{(i)}(h_i(x,y))})
  \end{split}
\end{equation}

In a multi-domain context, there are 2 ways to train domain-specific NMT model using adapters including pretraining generic NMT then finetuning adapters; or training both NMT model and adapters at once from the beginning using multi-task learning. In the literature of Machine Learning, \cite{Caruana97multitask} proposed an architecture suitable for multi-task learning. The corresponding network comprises a ground layer, which is shared between tasks and is followed by task-related layers. The network is then trained from scratch by fueling task-related examples in a batch and updating all parameters (shared and task-specific) at the same time. The multi-domain context is similar to the multi-task context if we consider a domain as a task\cite{Dredze08online}. Therefore, we can also use multi-task training to optimize the parameters of NMT model and the parameters of the residual adapters at once from the beginning. To assess the efficacy of multi-task training in the multi-domain context, we would like to compare the finetuning approach with the multi-task training approach. \fyTodo{Change the training regime}

\subsection{Highway Residual Adapters \label{ssec:highway}}
Finally, we would like to assess the performance of a variant of residual adapters in which the adapters go directly to the end of the stack of attention layers. We graphically demonstrate this version in Figure~\ref{fig:hrl-architecture}. In this version, only the last hidden vector of the encoder (decoder) is modified as follow
\begin{equation}
\bar{h}^n = h^n + \displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP^i(h^i)}
\end{equation}
The benefit of this variant is that it allows us to reuse hidden vectors $h^i$ of all hidden layers while adapting to multiple domains. In the forward step, the hidden vectors $h^i$ are calculated once, then adapters take these vectors to modify the last layer. Therefore, we can finetune the model to multiple domains at once without recomputing $h^i$. The second minor benefit of this variant is that it allows implementing the Gated Residual Adaptaters version that will be described in the following section \ref{ssec:gate}.
\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{fig/highway_residual}
  \caption{Highway multi-domain network with residual layers}
\label{fig:hrl-architecture}
\end{figure}
\subsection{Gated Residual Adaptaters \label{ssec:gate}}
\mpTodo{Formalizing problem, network design, training algorithm}
To translate a sentence, we have 2 options: using a domain-adapted model or using the generic model. In the case of choosing the wrong domain, the performance is dramatically worse than using the generic model. Therefore, we would like to design a version of applying residual adapter that allows us to automatically accept or reject given domain information, i.e, even if we choose a wrong adapter, the NMT model is still able to reject the change output from the adapter. In this version, we reuse the highway version of residual adapters. 
\fyTodo{Consistency of notations wrt section 2.1}
Given domain $k$ with set of adapters $ADAP^i_k$, in the model adapted to domain k, instead of using the formulation $\bar{h}^n = h^n + \displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP_k^i(h^i)}$ (encoder or decoder's side), we propose "gated" version uses $\bar{h}^n = h^n + \displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP_k^i(h^i) \odot{} z_k(h^n)}$. Where $z_k(h^n(w))$ is designed to reflect the relatedness of word $w$ to domain $k$: the more likely $w$ is in domain $k$, the larger $z_k(h^n(w))$ should be. The gated version aims to reduce the change $\displaystyle{\mathop{\sum}_{1 \leq i \leq n} ADAP_k^i(h^i)}$ added to $h^n$ for words that are not typical of domain $k$. In our implementation, we incorporate 2 domain classifiers on top of the encoder and the decoder that take the last hidden layer of encoder (decoder) as input, we take the posterior probability $P(d=k|w)$ as value of $z_k(h^n(w))$.

The training of Gated Residual Adaptaters comprises three main steps:
\begin{itemize}
	\item Pretraining a generic model with mixed corpora comprising data from multiple domains.
	\item Training a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen. This model computes the posterior domain probability $p(k|h(w_t))$ for each word $w_t$ based on the representation computed by the last layer.
	\item Training the parameters of adapters with in-domain data separately for each domain while freezing the parameters of the generic model and of the domain classifiers.
\end{itemize}

@@

\begin{table*}[htbp]
  \centering
  \begin{tabular}{ |lllllll|} %*{4}{|r|}}
    \hline
    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
    \domain{bank} & \domain{eco} & \domain{med} & \domain{gouv} & \domain{news} & \domain{tourism} & \domain{other} \\
    \hline
    4(0.00022) & 2857(0.15) & 347(0.018) & 1828(0.095) & 3696(0.19) & 7(0.00039) & 10473(0.54) \\
    \hline
  \end{tabular}
\caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture. \domain{news} is the largest domain, containing almost 19\% of the sentences, while \domain{bank} is the smallest, with only 0.02\% of the data.}
\label{tab:Corpora-en-de}
\end{table*}

@@

For all models in En-Fr, we set the embeddings size and the hidden layers size to~512. Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells. The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of a 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024. 
% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
Training use a batch size of~12288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}

For all models in En-De, we set the embeddings size and the hidden layers size to~1024. Transformers use multi-head attention with 16 heads in each of the 6 layers; the inner feedforward layer contains 4096 cells. The multi-domain residual system (see description below) additionally uses an adaptation block in each transformer layer, composed of a 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 2048. Training uses a batch size of~32768 tokens.

The size of models are reported in the table \ref{tab:params}

\begin{table*}[htbp]
  \centering
  \begin{tabular}{|l|l|} \hline
    Model & params \\ \hline 
    \system{Transformer-En-Fr}  & 65 \\
    \system{Residual Adapter-En-Fr} & 1 \\
    \system{Transformer-En-De}  & 213 \\
    \system{Residual Adapter-En-De} & 8 \\
     \hline
  \end{tabular}
  \caption{Number of parameters ($\times 10^6$)}
  \label{tab:params}
\end{table*}
\subsection{Varying the positions and number of residual adapters}
Tables \ref{tab:Corpora-en-fr} and \ref{tab:performance-en-de} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gouv}, \domain{eco}, \domain{tourism}, \domain{bank}, \domain{med} and \domain{news} for En-De. The original version \system{FT-Block} outperforms the baseline \system{mixed} in almost all domains while stays equivalent to the baseline in \domain{med} in En-Fr and \domain{news}, \domain{gouv}, \domain{eco}, \domain{med} in En-De. However, \system{FT-full} still stay as best model in all domain tests.

By varying the number and positions of residual adapters as aforementioned in section \ref{ssec:architecture}, we assess several variants of applying residual adapters. \fyTodo{Fix style here} Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 on each side of NMT model (encoder and decoder) which covers a wide range in the hierarchy of attention layers. To observe how much the number of residual adapters affects the performance of the adapted model. By the limit of computation resources, we limit our study to the case of 3 adapters at positions: 2,4,6 in each side of the NMT model. By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting) in each side (encoder and decoder).\fyTodo{We have 12 layers, have we not ?}

 In most cases, the performance increases with respect to the number of residual adapters used in architecture. The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level. However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?}

\begin{table*}
  \centering
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}} 
    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
    \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
    \system{FT-HW-Block}   & 37.5 & 57.2 & 53.4 & 33.1 & 46.3 & 91 & 53.1 \\ 
    \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 \\
    \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 \\
    \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 \\
   \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 \\
     \hline
  \end{tabular}
  \caption{Translation performance of various finetuned systems (En:Fr). We report BLEU scores for each domain, as well as averages.}\fyTodo{Boldface ?}
  \label{tab:performance-en-fr}
\end{table*}

@@

\fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ? It helps for small domain - domain-specific regularization ??}
\subsection{Regularization of fine-tuning}
In the translation from English into German, we find that domains \domain{tourism}, \domain{ecb} are extremely small and account for a very small percentage of the training data, with respective proportions of 0.039\% and 0.022\%. Fine-tuning on these domains can lead to serious overfitting. We assess two well-known regularization techniques for adapter modules, that could help avoid the overfitting issues: weight decay and layer regularization.

We choose optimal hyper-parameter $\lambda$ by grid search with a small set of values $\{ 10^{-3}, 10^{-4}, 10^{-5}$.

Results in Table~\ref{tab:performance-en-de-reg} show that weight decay (\system{FT-Block-WD}) outperforms the more basic \system{FT-Block}. \fyTodo{How is the weight decay parameter set ?}

\iffalse{
\begin{table*}[htbp]
  \centering
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}} 
    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 \\
   \system{FT-Block}     & 37.3	& 57.93 &	53.91 &	33.79 &	46.69 &	90.17 & 53.3 \\ 
   \system{FT-Block-WD}     & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 \\
   \system{FT-Block-LR}     & 37.45 & 56.09 & 51.84 & 33.29 & 45.02 & 89.7 & 52.2 \\
     \hline
  \end{tabular}
  \caption{Translation performance of various finetuned systems. We report BLEU scores for each domain, as well as averages.}
  \label{tab:performance-en-fr-reg}
\end{table*}
}
\fi

@@

\subsection{Multi-task learning}
To assess the effectiveness of multi-task learning in training both the NMT model and the residual adapters at once from scratch, we contrast our approach with several proposals from the literature on multi-domain MT including:
